---
title: "C++ Resources"
author: "Andrea Havron<br>Christine Stawitz<br>NOAA Fisheries, OST"
output:
  xaringan::moon_reader:
    self_contained: true
    css: ["xaringan-themer.css", "slides-style.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc( (%current% - 1) / (%total% - 1) * 100%);">
          </div>
        </div>`
---
layout: true

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]


<style type="text/css">

code.cpp{
  font-size: 14px;
}
code.r{
  font-size: 14px;
}


</style>


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-tile-view, echo=FALSE}
# this gives you a tile navigation if you type "O" at any time
xaringanExtra::use_tile_view()
```

---
# C++ and Memory


.column-60[
C++ code
```{Rcpp, eval = FALSE}
#include <iostream>

int main() {
  //initiate a variable
  float a = 3.1459;
  //initiate a new variable with the same value as a
  float b = a;
  //initiate a variable c that points to the same address as a
  float* c = &a;
  std::cout << "c is equal to the address of a; *c = a" << std::endl;
  std::cout << "c = " << c << std::endl;
  std::cout << "a = " << a << "; *c = " << *c << std::endl;
  return 0;
}
```
]

.column-40-left[
Console output
```{r, eval = FALSE}
g++ memoryexample.cpp -o a.exe
a.exe

c is equal to the address of a; *c = a
c = 0x78fe14
a = 3.1459; *c = 3.1459
```
]

---
# C++ and Memory

.column-60[
C++ code
```{Rcpp, eval = FALSE}
#include <iostream>

int main() {
  //initiate a variable
  float a = 3.1459;
  //initiate a new variable with the same value as a
  float b = a;
  //initiate a variable c that points to the same address as a
  float* c = &a;
  *c = 100;
  std::cout << "a and *c have been updated; b has not"  << std::endl;
  std::cout << "a = " << a << "; *c = " << *c << std::endl;
  std::cout << "b = " << b << std::endl;
  return 0;
}
```
]

.column-40-left[
Console output
```{r, eval = FALSE}
g++ memoryexample.cpp -o a.exe
a.exe

a and *c have been updated; b has not
a = 100; *c = 100
b = 3.1459
```
]
---
# C++ and Memory

.column-60[
C++ code
```{Rcpp, eval = FALSE}
#include <iostream>

int main() {
  //initiate a variable
  float a = 3.1459;
  //initiate a new variable with the same value as a
  float b = a;
  //initiate a variable c that points to the same address as a
  float* c = &a;
  *c = 100;
  c = &b;
  b = 10;
  a = 3;
  std::cout << "c now equals the address of b" << std::endl;
  std::cout << "c = " << c << "; &b = " << &b << std::endl;
  std::cout << "a = " << a << std::endl;
  std::cout << "b = " << b << std::endl; 
  std::cout << "*c = " << *c ;
  return 0;
}
```
]

.column-40-left[
Console output
```{r, eval = FALSE}
g++ memoryexample.cpp -o a.exe
a.exe

c now equals the address of b
c = 0x78fe10; &b = 0x78fe10
a = 3
b = 10
 *c = 10
```
]
---
# C++ Templates

.column-60[
C++ code
```{Rcpp, eval = FALSE}
#include <iostream>
#include <cmath>

template <class T>
  T my_add(T *x, T y){
    *x += y;
    return 0;
}
template <class T>
  T my_exp(T *x){
    *x = std::exp(*x);
    return 0;
}
int main() {
  //initiate a variable
  float par = 0;
  //initiate a new constant used to update par
  float b = 0.6931472;
  //intiate a pointer to par
  float *c = &par;
  //update par using functions
  my_add(c, b);
  std::cout << "par = " << par << std::endl;
  my_exp(c);
  std::cout << "par = " << par << std::endl;
  return 0;
}
```
]

.column-40-left[
Console output
```{r, eval = FALSE}
g++ pointerexample.cpp -o a.exe
a.exe

par = 0.6931472
par = 2
```
]

---
# C++ Anatomy of a Class
.column-60[
```cpp
class ClassName{
  //Access specifier: can be private, public, or protected
  public:
  //Data members: variables to be used
  float a;
  float b;
  
  ClassName(){} //Constructor - automatically called when new object created
  
  //Member functions: Methods to access data members
  my_add(){
    return a + b;
  }
  
}; //class name ends with semicolon


//create instance of class
ClassName m();
ClassName *m = new ClassName();

std::make_shared m //used in FIMS to create shared pointers to FIMS classes
```
<br>
resources:<br>
[**geeksforgeeks**](https://www.geeksforgeeks.org/c-classes-and-objects/)<br>
[**cplusplus**](https://cplusplus.com/doc/tutorial/classes/)
]

.column-40-left[
- Memory is not allocated when a class is defined
- An instance of the class needs to be created for memory allocation (new object created)
- **new** is the C++ initialize function that allocates memory (creates pointer)
- Classes can also be initialized as a declaration
- Once the class is instantiated, members can be accessed using
 - the dot(**.**) when class is declared
 - the pointer(**->**) when an instance of the class is created using **new**
 ]


---
# Class Example using templates

C++ code
.pull-left[
```{Rcpp, eval = FALSE}
#include <iostream>

template <class T>
class MyClass{
    public:
    T a;
    T b;
    MyClass(){}
    T evaluate(){
        return a + b;
    }
};
int main() {
  //initiate class with pointer
  MyClass<double> *m = new MyClass<double>();
  m -> a = 1.2;
  m -> b = 3.4;
  std::cout << m -> evaluate() << std::endl;  
  //intiate class with declaration
  MyClass<double> myclass;
  myclass.a = 2;
  myclass.b = 3;
  std::cout << myclass.evaluate() << std::endl;
  return 0;
}
```
]
.pull-right[
Console output
```{r, eval = FALSE}
g++ classexample.cpp -o a.exe
a.exe

4.6
5
```
]
---
# C++ Polymorphism

 - a function or operator works different when used in different context
 - virtual: C++ function that gets overridden at runtime
 - void: a C++ function with no return value
 
<br>

 .three-column-left[
```cpp
//include <iostream>

class Shape {
public:
  Shape();
  // Virtual function to calculate area
  virtual double area() {
    return 0.0; 
  }
};
```
 <br>
resources: <br>
[**geeksforgeeks**](https://www.geeksforgeeks.org/cpp-polymorphism/)<br>
[**cplusplus.com**](https://cplusplus.com/forum/general/275978/)
 ]

 .three-column-left[
```cpp
class Circle : public Shape {
private:
  double radius;

public:
  Circle(double r):Shape(){radius = r}

  // Override area() for Circle
  double area() {
    return M_PI * radius * radius;
  }
};
```
 ]

 .three-column-left[
```cpp
class Rectangle : public Shape {
private:
  double width, height;

public:
  Rectangle(double w, double h):Shape(){
  width = w;
  height = h;
}

  // Override area() for Rectangle
  double area()  {
    return width * height;
  }
};
```
 ]
---
# C++ Polymorphism

 .three-column-left[
```cpp
//include <iostream>

class Shape {
public:
  Shape();
  // Virtual function to calculate area
  virtual double area() {
    return 0.0; 
  }
};
```
 ]

 .three-column-left[
```cpp
class Circle : public Shape {
private:
  double radius;

public:
  Circle(double r):Shape(){radius = r}

  // Override area() for Circle
  double area() {
    return M_PI * radius * radius;
  }
};
```
 ]

 .three-column-left[
```cpp
class Rectangle : public Shape {
private:
  double width, height;

public:
  Rectangle(double w, double h):Shape(){
  width = w;
  height = h;
}

  // Override area() for Rectangle
  double area()  {
    return width * height;
  }
};
```
 ]
 
 <br>

.pull-left[
```cpp
int main()
{
Shape *a;
a = new Rectangle(4,5);
std::cout<<a->area()<<endl;

a = new Circle (2.4);
std::cout<<a->area()<<endl;
return 0;
}
```
]
.pull-right[
Console output
```{r, eval = FALSE}
g++ shape.cpp -o a.exe
a.exe

20
18.09557
```
]
 
---
class: middle

# Rcpp Interface
<br>
### Sources
- "Advanced `R`"  by Hadley Wickham (http://adv-r.had.co.nz/Rcpp.html)
- "`Rcpp`: Seamless `R` and `C++` integration" by Dirk Eddelbuttel and Romain Francois, Journal of Statistical Software (https://www.jstatsoft.org/article/view/v040i08)
- "`Rcpp` for everyone" by Masaki E. Tsuda (https://teuder.github.io/rcpp4everyone_en/)
- "`Rcpp` Modules" vignette by Eddelbuttel and Francois (https://cran.rstudio.com/web/packages/Rcpp/vignettes/Rcpp-modules.pdf)

---

# What is `RCpp` and why should I care?

.bluebox[## Definition
"The `RCpp` package provides a consistent API for seamlessly accessing, extending or modifying R objects at the C++ level" - Eddelbuttel and Francois.
]

.yellowbox[## API
a set of functions and procedures allowing the creation of applications that access the features or data of an operating system, application, or other service.]

* One of the most widely used `R` extensions (over 1000 packages)
* With very minimal knowledge of `C++` you can speed up function runs A LOT
- The most efficient `R` functions are written in `C++` and called from `R`


---
# Writing `C++` functions in `R`

* You can write `C++` functions inline in your R code using `cppFunction`
* You can also compile single lines of code directly using `evalCpp`

In R script or console:
```{r}
library(Rcpp)
cppFunction('int add(int x, int y, int z) {
  int sum = x + y + z;
  return sum;
}')
# add works like a regular R function
add
add(1, 2, 3)
evalCpp("std::numeric_limits<double>::max()")
```

---
# Calling `C++` from `R`

.pull-left[
mean.cpp
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double meanC(NumericVector x) {
  int n = x.size();
  double total = 0;

  for(int i = 0; i < n; ++i) {
    total += x[i];
  }
  return total / n;
}

```
]

.pull-right[
* `#include <Rcpp.h>` allows you to call `Rcpp` syntax in the file. You can omit `using namespace Rcpp` if you use `::` i.e. `Rcpp::NumericVector` 
* `// [[Rcpp::export]]` - put this above any function you want to make available in `R`
* `NumericVector, .size()` useful type, function implemented in `Rcpp`]

---
# Calling `C++` from `R`
.pull-left[
```{r eval=FALSE}
sourceCpp("mean.cpp")
```


```{r}
library(microbenchmark)
x <- runif(1e5)
microbenchmark(
  mean(x),
  meanC(x)
)

```
]


.pull-right[
* `sourceCpp` compiles the file
* More `Rcpp` syntax [here](https://dirk.eddelbuettel.com/code/rcpp/Rcpp-quickref.pdf)
]

---

# Typing and methods
.pull-left-narrow[
- Scalar classes
* `int`
* `double`
* `bool`
* `String`
- vector classes
* `IntegerVector`
* `NumericVector`
* `LogicalVector`
* `CharacterVector`
]

.pull-right-wide[
- Vector classes have useful static methods (called with `::` on class) 
  ```{Rcpp eval=FALSE}
   NumericVector v = NumericVector::create(1,2,3);
  NumericVector::get_NA(); 
  ```
- Member functions or methods (called with `.` on object)
  ```{Rcpp eval=FALSE} 
  v.size(); 
  v.push_back(4);
  ```
  operate on object `v`
]

---
# `Rcpp` modules


```{Rcpp eval=FALSE}
RCPP_MODULE(unif_module) {
class_<Uniform>("Uniform")
.constructor<double,double>()
.field("min", &Uniform::min, "minimum value")
.field("max", &Uniform::max, "maximum value")
.method("draw", &Uniform::draw)
;
}
```

* You can expose these using the `Rcpp` macro `RCPP_MODULE`
* the `class_<Uniform>` constructor argument is what we will call the class from `R`
* `.field`, `.constructor`, `.method`, `.property` 
* `.field` can be used with two or three arguments
* `.field_readonly` prevents it from being modified from `R` 

