---
title: "Cpp Inheritance"
author: "Andrea Havron"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## C++ Inheritence

- When a new class (child, derived) inherits members and functions from an existing class (parent, base)
- There are three different types of inheritence:
    - **Public**: everything that can access the base class can access its public members and functions
    - **Protected**: only the derived classes can access the protected members and functions of a base class
    - **Private**: only the base class can access private members and functions of the base class
- FIMS only uses Public inhertience so we will focus on those examples
- **this->** used to reference the base class or local members within the derived class

When setting up inheritance, its best to establish a base class with behavior shared with all the derived classes.

```cpp
class Shape{

public:
double length;
double height;


Shape(){}

};

```

Then derived classes can inherit from the base class

```cpp
class Rectangle : public Shape{
public:  
  Rectangle() : Shape() {}

  public:
  //this->length: points to length within base class, etc.
  double area(){
    return this->length * this->height;
  }
};
  
class Triangle : public Shape{
public:
  Triangle() : Shape() {}
  
  public:
  double area(){
    return 0.5 * this->length * this->height;
  } 
};
```

The Rcpp function can then calculate the areas of the different shapes:
```cpp
// [[Rcpp::export]]
double calculate_areas(std::string shape, double length, double height){
  
  double out = 0;
  
  if(shape == "rectangle"){
      Rectangle rect;
      rect.length = length;
      rect.height = height;
      out = rect.area();
  } else if (shape == "triangle") {
      Triangle tri;
      tri.length = length;
      tri.height = height;
      out = tri.area();
  } else {
      Rcpp::Rcout << "Invalid shape" << std::endl;
  }
  return out;
}
```

All the code is combined into `my_inheritance.cpp` in the `src` directory and can be run from r using `Rcpp::sourceCpp`
```{r run-inheritance}
Rcpp::sourceCpp("../src/my_inheritance.cpp")

calculate_areas("triangle", 6, 4)
calculate_areas("square", 2, 2)
```


