---
title:  "population.hpp review"
author: "Andrea Havron<br>NOAA Fisheries, OST"
output:
  xaringan::moon_reader:
    self_contained: true
    css: ["xaringan-themer.css", "slides-style.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc( (%current% - 1) / (%total% - 1) * 100%);">
          </div>
        </div>`
---
layout: true

.footnote[U.S. Department of Commerce | National Oceanic and Atmospheric Administration | National Marine Fisheries Service]


<style type="text/css">

code.cpp{
  font-size: 14px;
}
code.r{
  font-size: 14px;
}


</style>


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-tile-view, echo=FALSE}
# this gives you a tile navigation if you type "O" at any time
xaringanExtra::use_tile_view()
```

---
# C++ Anatomy of a Class

```cpp
class ClassName{
  Access specifier: //can be private, public, or protected
  Data members; //variables to be used
  
  ClassName(){} //Constructor - automatically called when new object created
  
  Member functions(){} //Methods to access data members
  
}; //class name ends with semicolon


// create pointer to new class object
ClassName<T>* ClassName<T>::instance = new ClassName<T>();
// or declare new class object
ClassName<T> myclass;
```
- Memory is not allocated when a class is defined
- An instance of the class needs to be created for memory allocation (new object created)
- **new** is the C++ initialize function that allocates memory
- Class members can be accessed using the dot('.') if using an object or arrow('->') if using a pointer
<br><br>
Resources:<br>
[**geeksforgeeks**](https://www.geeksforgeeks.org/c-classes-and-objects/)<br>
[**cplusplus**](https://cplusplus.com/doc/tutorial/classes/)
---
# Class Example

C++ code
.pull-left[
```{Rcpp, eval = FALSE}
#include <iostream>

template <class T>
class MyClass{
    public:
    T a;
    T b;
    MyClass(){}
    T evaluate(){
        return a + b;
    }
};
int main() {
  //initiate class
  MyClass<double> *m = new MyClass<double>();
  m -> a = 1.2;
  m -> b = 3.4;
  std::cout << m -> evaluate() << std::endl;  
  //intiate class
  MyClass<double> myclass;
  myclass.a = 2;
  myclass.b = 3;
  std::cout << myclass.evaluate() << std::endl;
  return 0;
}
```
]
.pull-right[
Console output
```{r, eval = FALSE}
g++ classexample.cpp -o a.exe
a.exe

4.6
5
```
]
---
# population.hpp
```cpp
template <typename Type>
  struct Population : public FIMSObject<Type>{
  
  //Data members: variables to be used
  static uint32_t id_g; /*!< reference id for population object*/
  size_t nyears; /*!< total number of years in the fishery*/
  ...
  // constants
  // parameters
  // Transformed values
  // derived quantities
  
  /// recruitment
    int recruitment_id = -999;                                /*!< id of recruitment model object*/
    std::shared_ptr<fims::RecruitmentBase<Type>> recruitment; /*!< shared pointer to recruitment module */

  Population()
  {
    this->id = Population::id_g++;
  }
  
  Member functions(){} //Methods to access data members
  
}; //class name ends with semicolon

template <class Type>
uint32_t Population<Type>::id_g = 0;

```
---
# Member Functions

```cpp

    /**
     * @brief Initialize values. Called once at the start of model run.
     *
     * @param nyears total number of years in the fishery
     * @param nseasons total number of seasons in the fishery
     * @param nages total number of ages in the population
     */
    void Initialize(int nyears, int nseasons, int nages){...}
    
   /**
     * @brief Prepare to run the population loop. Called at each model itartion, and used
     *  to zero out derived quantities, values that were summed, etc.
     *
     */
    void Prepare(){...}
    
    void CalculateMaturityAA(int index_ya, int age)
    {
      // this->maturity is pointing to the maturity module, which has
      //  an evaluate function. -> can be nested.
      this->proportion_mature_at_age[index_ya] =
          this->maturity->evaluate(age);
    }
```
---
# Additional Member Functions

<br>
```cpp
CalculateInitialNumbersAA(...){...}

CalculateMortality(...){...}

CalculateNumbersAA(...){...}

CalculateUnfishedNumbersAA(...){...}

CalculateSpawningBiomass(...){...}

CalculateUnfishedSpawningBiomass(...){...}

CalculateRecruitment(...){...}

CalculateCatch(...){...}

CalculateIndex(...){...}

CalculateCatchNumbersAA(...){...}

CalculateCatchWeightAA(...){...}

Evaluate(...){...}
```
---
class: middle

[github.com/FIMS/180-feat-population/inst/include/population_dynamics/population/](https://github.com/NOAA-FIMS/FIMS/blob/180-feat-population/inst/include/population_dynamics/population/population.hpp)

